---
title: "Gravity Models"
output: html_document
editor_options: 
  chunk_output_type: console
---

Load libraries.
```{r message=FALSE, warning=FALSE}
library(caret)
library(rpart)
library(randomForest)
library(kernlab)
library(nnet)
library(NeuralNetTools)
```

Read data.
```{r}
trainData <- read.table("gravityTrainData.csv", sep=",", header=FALSE)
testData <- read.table("gravityTestData.csv", sep=",", header=FALSE)
colnames(trainData) <- c("F", "m1", "m2", "r")
colnames(testData) <- c("F", "m1", "m2", "r")
```

Calculate the gravitational constant.
```{r}
summary(gravConst <- trainData$F/(trainData$m1*trainData$m2/trainData$r^2))
```

Linear regression with main effects.  All terms are significant.
```{r}
mylm <- lm(F ~ ., data=trainData)
summary(mylm)

mylmPredict <- predict(mylm, newdata=testData)

(lm_pr <- postResample(pred = mylmPredict, obs = testData$F))
```

Fit a second-order model.  All interaction terms are significant, as well as $m_1^2$.  
```{r}
mylmIntAll <- lm(F ~ (.)^2 + I(m1^2) + I(m2^2) + I(r^2), data=trainData)
summary(mylmIntAll)

mylmIntAllPredict <- predict(mylmIntAll, newdata=testData)

(lm_int_all_pr <- postResample(pred = mylmIntAllPredict, obs = testData$F))
```

```{r warning=FALSE}
myrpart <- train(x = trainData[,-1],
                y = trainData[,1],
                method="rpart")

plot(myrpart$finalModel)
text(myrpart$finalModel)

myrpart$finalModel$variable.importance

myrpartPredict <- predict(myrpart, newdata=testData[,-1])
(rpart_pr <- postResample(pred = myrpartPredict, obs = testData$F))
```

Support vector machine with a linear kernel.
```{r}
mysvmLinear <- train(x = trainData[,-1],
                     y = trainData[,1],
                     method="svmLinear")

mysvmLinearPredict <- predict(mysvmLinear, newdata=testData[,-1])
(svm_lin_pr <- postResample(pred = mysvmLinearPredict, obs = testData$F))
```

Support vector machine with a polynomial kernel.
```{r}
mysvmPoly <- train(x = trainData[,-1],
                     y = trainData[,1],
                     method="svmPoly")

mysvmPolyPredict <- predict(mysvmPoly, newdata=testData[,-1])
(svm_poly_pr <- postResample(pred = mysvmPolyPredict, obs = testData$F))
```

Support vector machine with a radial basis function kernel.
```{r}
mysvmRbf <- train(x = trainData[,-1],
                     y = trainData[,1],
                     method="svmRadial")

mysvmRbfPredict <- predict(mysvmRbf, newdata=testData[,-1])
(svm_rbf_pr <- postResample(pred = mysvmRbfPredict, obs = testData$F))
```

Neural network.
```{r message=FALSE}
mynnet <- train(x = trainData[,-1],
                y = trainData[,1],
                method="nnet",
                trace=FALSE)

garson(mynnet$finalModel, bar_plot=FALSE)

mynnetPredict <- predict(mynnet, newdata=testData[,-1])
(nnet_pr <- postResample(pred = mynnetPredict, obs = testData$F))
```

Random forest.
```{r}
myrf <- train(x = trainData[,-1],
              y = trainData[,1],
              method="rf")

myrf$finalModel$importance


myrfPredict <- predict(myrf, newdata=testData[,-1])
(rf_pr <- postResample(pred = myrfPredict, obs = testData$F))
```

RMSE plot.
```{r}
rmse_data <- data.frame(
  model=c("LM", "LM_INT", "RT", "SVM_LIN", "SVM_POLY",  "SVM_RBF", "NN", "RF"),
  RMSE=c(lm_pr[1], lm_int_all_pr[1], 
         rpart_pr[1], svm_lin_pr[1],
         svm_poly_pr[1], svm_rbf_pr[1],
         nnet_pr[1], rf_pr[1])
)
rmse_data$model <- factor(rmse_data$model, levels=c("LM", "LM_INT", "RT", "SVM_LIN", "SVM_POLY",  "SVM_RBF", "NN", "RF"))


ggplot(rmse_data, aes(x=model, y=RMSE)) + 
  geom_bar(stat="identity", color="black", fill="blue") +
  theme(axis.text.x=element_text(angle=90, hjust=1))
```


# CONJECTURING


Recover the gravitational constant using linear regression, based on the relationship $F \geq m_1m_2/r^2$.

```{r}
constDF <- data.frame(F=trainData$F, m1m2OverR2 = trainData$m1*trainData$m2/trainData$r^2)

constlm <- lm(F ~ ., data=constDF)
summary(constlm)
coef(constlm)
```


